---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LBBNN

<!-- badges: start -->

[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![](https://img.shields.io/github/last-commit/LarsELund/LBBNN.svg)](https://github.com/LarsELund/LBBNN/commits/main) [![](https://img.shields.io/github/languages/code-size/LarsELund/LBBNN.svg)](https://github.com/LarsELund/LBBNN) [![R-CMD-check](https://github.com/LarsELund/LBBNN/workflows/R-CMD-check/badge.svg)](https://github.com/LarsELund/LBBNN/actions) [![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE)

<!-- badges: end -->

The package implements LBBNN (Latent Binary Bayesian Neural Networks) (<https://openreview.net/pdf?id=d6kqUKzG3V>) in R using the LibTorch backend via the Torch package. Currently, standard LBBNNs are implemented, using the local reparametrization tricks and optionally normalizing flows. In addition, the input-skip architecture is also implemented. <https://arxiv.org/abs/2503.10496>).

## Installation

You can install the development version of LBBNN from [GitHub](https://github.com/LarsELund/LBBNN) with:

``` r
# install.packages("pak")
pak::pak("LarsELund/LBBNN")
```

## Example

This example demonstrates how to implement a simple feed forward LBBNN on a small dataset
containing morphological features of two types of raisins. 

The get_dataloaders() function takes a data.frame object, and transforms it to 
a tensor_dataset, which can be used in a torch_dataloader object. This enables 
automatic mini-batching and parallel data loading. 

The arguments define the proportion of data to be used for training vs validation
in addition to the batch sizes for the respective dataloaders.

```{r example}
library(LBBNN)
loaders <- get_dataloaders(raisin_dataset, train_proportion = 0.8, 
                           train_batch_size = 720, test_batch_size = 180)
train_loader <- loaders$train_loader
test_loader <- loaders$test_loader

```

Important hyperparameters include the size of the network, priors, and initializations.
When defining an 'lbbnn_net' object, the sizes argument is a vector of integers, where the 
first entry represents the number of features (7 in this case), the subsequent entries
are the widths of the hidden layers, and the last entry the number of outputs.

The prior argument defines the inclusion priors for each layer, whereas std refers
to the standard deviation of the weight priors. inclusion_inits refers to the 
initialization of the variationa linclusion parameters. It controls the initial
density of the network.

```{r}
problem <- 'binary classification'
sizes <- c(7, 5, 5, 1) 
inclusion_priors <-c(0.5, 0.5, 0.5) 
stds <- c(1, 1, 1) 
inclusion_inits <- matrix(rep(c(-10, 15), 3),nrow = 2, ncol = 3)
device <- 'cpu' #can also be mps or gpu.
```

Below we define the model. We use the input_skip architecture in this case, but not normalizing flows.

```{r}
torch::torch_manual_seed(0)
model_input_skip <- lbbnn_net(problem_type = problem,sizes = sizes,prior = inclusion_priors,
                      inclusion_inits = inclusion_inits,input_skip = TRUE,std = stds,
                   flow = FALSE,device = device)
```

The function lbbnn_train is used to train the model. In this case for 800 epochs with a learning rate of 0.01.
During training, the loss and accuracy can be displayed at each epoch.

```{r,results='hide'}
results_input_skip <- suppressMessages(train_lbbnn(epochs = 800,LBBNN = model_input_skip, lr = 0.01,train_dl = train_loader,device = device))
#save the model 
#torch::torch_save(model_input_skip$state_dict(), 
#paste(getwd(),'/R/saved_models/README_input_skip_example_model.pth',sep = ''))
```

Below, validate_lbbnn is used to test the model on unseen data. We provide
the test_loader, containing the data set aside.
```{r}
validate_lbbnn(LBBNN = model_input_skip,num_samples = 100,test_dl = test_loader,device)
#validate_lbbnn(LBBNN = model_flows,num_samples = 1000,test_dl = test_loader,device)
```

The package provides utilities for both global and local explanations. Global explanations
describe how the models behaves overall across the entire dataset. The function plot, 
with argument type = 'global' visualizes the structure of the network, when only weights within
active paths are included, where an active path is a path from an input, to the output, either directly
or through one or more hidden neurons.

```{r}
plot(model_input_skip,type = 'global',vertex_size = 13,edge_width = 0.6,label_size = 0.6)
```

We see that only 4 of the 7 input variables are included..

The summary function provides further detail into inclusion probabilities from different layers.

```{r}
summary(model_input_skip)
```

Local explanations aim to explain a model’s prediction for a specific observation by quantifying how each input feature influences that prediction. Below we show one example. Unlike other methods like SHAP and LIME, our local explanations are intrinsic, and we also get uncertainty around the explanations.
```{r}
x <- train_loader$dataset$tensors[[1]] #grab the dataset
ind <- 42
data <- x[ind,] #plot this specific data-point
plot(model_input_skip,type = 'local',data = data,num_samples = 100)
```

The residual function computes the residuals: y_true - y_predicted

```{r}
residuals(model_input_skip)[1:10]
```

coef gives the local explanations average over multiple samples:

```{r}
coef(model_input_skip,dataset = train_loader,inds = c(2,3,4,5,6))
```

posterior predictions:

```{r}
predictions <- predict(model_input_skip,mpm = TRUE,newdata = test_loader,draws = 100)
dim(predictions) #shape is (draws,samples,classes)
```

Print the model:

```{r}
print(model_input_skip)
```
