---
title: "Classification on a small dataset"
output: rmarkdown::html_vignette
params:
  eval: true
vignette: >
  %\VignetteIndexEntry{classification: gallstone dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prepare data

We use the gallstone dataset, consisting of a mix of demographic, bioimpedance and laboratory features. It is a mix of binary and continuous features, with 40 in total after pre-processing the dataset. It is relatively small, with only 319 total samples, where we use 223 of those for training and the remaining 96 for validation.
```{r}
library(LBBNN)
torch::torch_manual_seed(42)
loaders_gs <- get_dataloaders(gallstone_dataset, train_proportion = 0.70,
                              train_batch_size = 223, test_batch_size = 96,
                              standardize = TRUE, seed = 42)
train_loader_gs <- loaders_gs$train_loader
test_loader_gs <- loaders_gs$test_loader
```

## Define hyperparameters and model object

A relatively large network consisting of 4 hidden layers with 16 neurons each is used in this example.  We also use normalizing flows for this example. We actually have over 30000 model parameters here when including all the flow parameters, which can be seen by using the print function.

```{r}
problem <- "binary classification"
sizes <- c(40, 16, 16, 16, 16, 1)
inclusion_priors <- c(0.5, 0.5, 0.5, 0.5, 0.5) 
stds <- c(1, 1, 1, 1, 1) 
inclusion_inits <- 'polarized_dense'
device <- "cpu"
model_gs <- lbbnn_net(problem_type = problem, sizes = sizes,
                      prior = inclusion_priors,
                      inclusion_inits = inclusion_inits,
                      input_skip = TRUE, std = stds, flow = TRUE,
                      dims = c(10, 10 ,10, 10), device = device)
#print(model_gs)
```

## Train and validate 

```{r}
results_gs <- train_lbbnn(epochs = 1000, LBBNN = model_gs,
                          lr = 0.005, train_dl = train_loader_gs,
                          device = device, scheduler = "step",
                          sch_step_size = 1000, verbose = FALSE)

validate_lbbnn(LBBNN = model_gs, num_samples = 100,
               test_dl = test_loader_gs, device)
```
We see that the accuracy is not great, especially on the sparse model. However, the density in active paths 
is only 0.006, perhaps too small to obtain good performance. What if we try a smaller architecture? The below model
has around 11000 parameters.

```{r}
torch::torch_manual_seed(42)
sizes <- c(40, 3, 3, 1)
inclusion_priors <- c(0.5, 0.5, 0.5) 
stds <- c(1, 1, 1) 
inclusion_inits <- 'polarized_dense'
device <- "cpu"
model_smaller <- lbbnn_net(problem_type = problem, sizes = sizes,
                      prior = inclusion_priors,
                      inclusion_inits = inclusion_inits,
                      input_skip = TRUE, std = stds, flow = TRUE,
                      dims = c(10, 10 ,10, 10), device = device)
```

```{r}
results_2 <- train_lbbnn(epochs = 1000, LBBNN = model_smaller,
                          lr = 0.005, train_dl = train_loader_gs,
                          device = device, scheduler = "step",
                          sch_step_size = 1000, verbose = FALSE)

validate_lbbnn(LBBNN = model_smaller, num_samples = 100,
               test_dl = test_loader_gs, device)
```
In this case, accuracy is much better even though we still have a very sparse model.

## Terminating training at a set density

To avoid structures that are too sparse, it is possible to use the argument 'min_density' in the train_lbbnn function to stop 
training when this density is reached. 

```{r}
torch::torch_manual_seed(42)
sizes <- c(40, 16, 16, 16, 16, 1)
inclusion_priors <- c(0.5, 0.5, 0.5, 0.5, 0.5) 
stds <- c(1, 1, 1, 1, 1) 
inclusion_inits <- 'polarized_dense'
device <- "cpu"
model_stopped <- lbbnn_net(problem_type = problem, sizes = sizes,
                      prior = inclusion_priors,
                      inclusion_inits = inclusion_inits,
                      input_skip = TRUE, std = stds, flow = TRUE,
                      dims = c(10, 10 ,10, 10), device = device)

results_3 <- train_lbbnn(epochs = 1000, LBBNN = model_stopped,
                          lr = 0.005, train_dl = train_loader_gs,
                          device = device, scheduler = "step",
                          sch_step_size = 1000, verbose = FALSE, min_density = 0.10)

validate_lbbnn(LBBNN = model_stopped, num_samples = 100,
               test_dl = test_loader_gs, device)

```
The accuracy is a little improved here. However, this dataset is in general quite sensitive to hyperparameters such as network architecture,
number of training epochs and parameter initializations.








