---
title: "variable selection: linear-data"
output: rmarkdown::html_vignette
params:
  eval: true
vignette: >
  %\VignetteIndexEntry{variable selection: linear-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Generate data

We generate 1000 samples with 15 features, and make 3 of them relevant for the outcome. The goal is to see if we recover the 
correct variables, and also their coefficients. The data preparation process follows the same steps as in the previous example.
Note that the dataset must be a data.frame object before applying the get_dataloaders function.
```{r}
library(LBBNN)
i <- 1000
j <- 15
set.seed(42)
torch::torch_manual_seed(42)
X <- matrix(rnorm(i * j, mean = 0, sd = 1), ncol = j)
y_base <- c()
y_base <-  0.6 * X[, 1] - 0.4 * X[, 2] + 0.5 * X[, 3] + rnorm(n = i, sd = 0.1)
sim_data <- as.data.frame(X)
sim_data <- cbind(sim_data, y_base)
loaders <- get_dataloaders(sim_data, train_proportion = 0.9,
                           train_batch_size = 450, test_batch_size = 100,
                           standardize = FALSE)
train_loader <- loaders$train_loader
test_loader  <- loaders$test_loader
```

## Define hyperparameters and the model object

Here we must use the "regression" keyword as we have a continuous outcome. We use 
a relatively small network of 2 hidden layers with 5 neurons in each. 
```{r}
problem <- "regression"
sizes <- c(j, 5, 5, 1) 
incl_priors <- c(0.5, 0.5, 0.5) 
stds <- c(1, 1, 1)
incl_inits <- 'polarized'
device <- "cpu" 
model_linear <- lbbnn_net(problem_type = problem, sizes = sizes,
                              prior = incl_priors, inclusion_inits = incl_inits,
                              std = stds, input_skip = TRUE, flow = FALSE,
                              num_transforms = 2, dims = c(10, 10, 10),
                              raw_output = FALSE, custom_act = NULL,
                              link = NULL, nll = NULL,
                              bias_inclusion_prob = FALSE, device = device)
```

## Train and validate the model

```{r}
train_lbbnn(epochs = 50, LBBNN = model_linear,
            lr = 0.1, train_dl = train_loader, device = device, verbose = FALSE)
validate_lbbnn(LBBNN = model_linear, num_samples = 2, test_dl = test_loader,
              device = device)
```

## Inspect the results using coef

Here we have given the dataset as an argument instead of picking a specific sample,
where the argument num_data controls how many samples from the dataset we use. Using inds = NULL,
we select random samples from the dataset. Explanations are then averaged across these samples.
```{r}
coef(model_linear, dataset = train_loader, inds = NULL,
     output_neuron = 1, num_data = 5, num_samples = 10)
```
We see that the coefficients match the data generating process, and that we only include the relevant features.

## Global structure

Using plot with type = 'global' gives us the global explanations, i.e. the features that affect predictions 
in general.
```{r,,fig.width=6, fig.height=6}
plot(model_linear, type = "global", vertex_size = 7,
     edge_width = 0.4, label_size = 0.4)
```

We see that everything has been pruned away aside from the linear connections from the relevant features.
